---
title: "OPP Veil of Darkness Report"
author: "Samantha Robertson"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load}
suppressMessages(source(here::here("lib", "opp.R")))
suppressMessages(source(here::here("lib", "veil_of_darkness_test.R")))

data.cities <- read_rds(here::here("cache", "vod_full_data.Rds"))
data.states <- read_rds(here::here("cache", "vod_statewide_full_data.Rds"))
```

# Introduction
To explore discrimination at the stop level, we implement the Veil of Darkness (VOD) test as proposed by Grogger & Ridgeway (G & R) in 2006. The test uses darkness as a proxy for the visibility of a driver’s race, and assumes that, in the presence of discrimination, a minority driver is more likely than a white driver to be stopped during the day when their race is visible. At night, the driver's race is assumed to be less visible, so the probability that a minority driver is stopped should be closer to the probability that a white driver is stopped.

# Summary of the dataset
We consider both city-level stops and statewide highway patrol stops. As in G & R, the analysis is limited to black and white drivers. Throughout the report, the term "minority driver" always refers to black drivers. The requirements of the test are quite stringent: the data must contain date, time and location of the stop, and the race of the driver. An additional control considered is geographic subdivision. The analysis is ultimately restricted to 7 cities and 10 states with all of these variables present.[^1]

[^1]: There are 26 cities with at least 75% non-`NA` values for date, time, driver race and stop location. 9 cities (Dallas, El Paso, Green Bay, Camden, Bakersfield, Little Rock, Plano, Hartford and Owensboro) are removed due to data issues, such as duplication, insufficient data in the intertwilight zone, or a negligible number of stops of minority drivers. Another 10 cities (Mesa, San Jose, Aurora, Wichita, Saint Paul, Grand Forks, Oklahoma City, Tulsa, San Antonio and Burlington) are removed because they do not provide one of region, precinct, reporting_area, district, beat, sector or police_grid_number. Though not included in this report due to time constraints, it may be interesting to run the regression that does not adjust for geographic subdivision on a larger dataset that includes these latter 10 cities.

# Methods

## Data Processing
As emphasised by Grogger & Ridgeway, racial driving patterns may vary significantly with time, so the analysis is limited to the “intertwilight zone”, which is the period of time during which it is light at least one day of the year and dark at least one day of the year. Since the darkness between sunset and dusk is ambiguous, “darkness” is defined as after dusk and “lightness” is defined as before sunset. Thus the intertwilight zone is defined as the minutes between the earliest dusk and the latest sunset. In addition, any stops between sunset and dusk are excluded due to the ambiguity.

A summary of the processed data included in the analysis is provided in Table 1 for the city level data and in Table 2 for the statewide data. Note that the number of stops and percentage of minority drivers are calculated after limiting the data to only stops in the intertwilight zone and only black and white drivers.

```{r data_summary_cities}
col_names <-
  c(
    "City",
    "Total Number of Stops", 
    "Year Range", 
    "Percentage of Minority Drivers"
  )

data.cities %>% 
  group_by(city) %>% 
  summarise(
    n_total = n(),
    min_year = min(year(date)), 
    max_year = max(year(date)),
    year_range = glue::glue("{min_year}-{max_year}"),
    percentage_minority = scales::percent(sum(is_minority_demographic) / n_total)
  ) %>% 
  select(-min_year, -max_year) %>% 
  arrange(desc(n_total)) %>% 
  knitr::kable(col.names = col_names, caption = "Summary of stops (city-level)")
```

```{r data_summary_states}
col_names <-
  c(
    "State",
    "Total Number of Stops", 
    "Year Range", 
    "Percentage of Minority Drivers"
  )

data.states %>% 
  group_by(state) %>% 
  summarise(
    n_total = n(),
    min_year = min(year(date)), 
    max_year = max(year(date)),
    year_range = glue::glue("{min_year}-{max_year}"),
    percentage_minority = scales::percent(sum(is_minority_demographic) / n_total)
  ) %>% 
  select(-min_year, -max_year) %>% 
  arrange(desc(n_total)) %>% 
  knitr::kable(caption = "Summary of stops (statewide)", col.names = col_names)
```


The calculation of sunset and dusk time requires latitude and longitude of the stop and the corresponding timezone. The raw data rarely includes this data, so latitude and longitude are estimated. For the city data, latitude and longitude are estimated using Google Maps, based on the officer’s description of the stop location. This procedure invariably makes a few errors, so any points more than 0.5 degrees in magnitude from the median latitude and longitude are considered spurious outliers and are dropped from the dataset. For the statewide data, the county centroid is used[^2].

[^2]: Shapefiles are used from the IPUMS NHGIS American Community Survey, 2016. Source: Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles. IPUMS National Historical Geographic Information System: Version 13.0 [Database]. Minneapolis: University of Minnesota. 2018. http://doi.org/10.18128/D050.V13.0

We define geographic subdivisions to be use as a control in the regression analyses. For the statewide data, county is the appropriate subdivision, however the most appropriate subdivision varies by city. The subdivision selected for each city, one of precinct, district, beat or sector, is the geographic subdivision variable that has between 5 and 30 distinct values available in the data.

Calculating sunset and dusk times is very slow, so the time of sunset and dusk is calculated only once for each geographic subdivision. For the cities, the sunset and dusk times are calculated at the median latitude and longitude of the stops in the subdivision. For the statewide data, these times are calculated at the county centroid. The intertwilight zone is then defined for each geographic subdivision. This avoids the possibility that stops are included in the analysis at a time and location when and where it is either always light or always dark.

## Analysis

To assess the extent of VOD discrimination, we estimate the change in the proportion of minority drivers stopped from when it is light to when it is dark using both visualizations and logistic regression.

There are a number of factors to consider that may affect the proportion of minority drivers in the at-risk population. Firstly, since driving patterns change over time, we control for clock time in all visualizations and all logistic regression models. As in G & R, we control for clock time in all models with the natural spline of the stop time with six degrees of freedom. To formalize the regression models, let $b$ be the proportion of minority drivers, $d$ be an indicator variable where $d = 1$ if it is dark and $d = 0$ otherwise, and $t$ be the stop time in minutes since midnight. The simplest model is then $d = \beta_0 + \beta_1d + \gamma_1^Tns_6(t)$.

In addition, we fit a second model that additionally controls for and geographic subdivision with an additive categorical variable. Since it is reasonable, if not likely, that discrimination varies greatly by city or state, we also fit this time and geography controlled model on each city and each state individually.

A concern raised by G & R is the role of seasonal effects. To control for seasonal effects we fit a third model on the full dataset with an additional additive control that treats the month of the year as a categorical variable. In addition, we fit the time and geography-controlled regression on subsets of the data that includes only the two weeks before and after the two annual Daylight Savings Time (DST) shifts.[^3]

[^3]: The model fit with a month control, and the DST models are only fit on the city-level data, due to time constraints and the large size of the aggregated statewide dataset, however the code is readily available to fit these models on the statewide data.

The DST shift provides a sharp discontinuity in sunset times that creates a period of one hour that is dark before the shift and light afterwards, or vice versa. The direction of the shift is opposite in the Fall and in the Spring, so we fit separate models for the two shifts. Since the data available to each model is limited to four consecutive weeks of the year, the intertwilight zone is redefined to limit the analysis to stops between the earliest dusk and the latest sunset in these four weeks. This significantly reduces the quantity of data. Table 3 summarizes the data available to the Fall and Spring DST models.

```{r summary_dst_cities}
data.dst_fall <- read_rds(here::here("cache", "vod_dst_fall_data.Rds"))
data.dst_spring <- read_rds(here::here("cache", "vod_dst_spring_data.Rds"))

data.dst_fall %>% 
  count(city, sort = TRUE) %>% 
  rename(n_fall = n) %>% 
  left_join(count(data.dst_spring, city), by = "city") %>% 
  rename(n_spring = n) %>% 
  knitr::kable(
    caption = "Summary of intertwilight zone stops in the Fall and Spring DST periods", 
    col.names = c(
      "City", 
      "Number of stops (Fall)",  
      "Number of stops (Spring)"
    )
  )
```

G & R define the veil of darkness parameter $K_{vod} = \frac{P(B|S, d = 0)P(\bar{B} | S, d = 1)}{P(\bar{B}|S, d = 0)P(B | S, d = 1)}$, where $B$ is the event a driver is black, $S$ is the event a driver is stopped, and $d$ is darkness as we define it, where $d = 1$ indicates it is dark and $d = 0$ indicates it is light. Theier analysis focuses on $log K$, which is negative in the presence of discrimination. As described in the paper, $-\beta_1$, where $\beta_1$ is the coefficient on the darkness variable in the logistic regression models, is an estimate for $logK$.  In this analysis we choose to consider the coefficient directly, as we find this a more intuitive and interpretable approach. A negative coefficient $\beta_1$ indicates that darkness is associated with a lower proportion of minorities stopped, and thus a negative coefficient is indicative of VOD discrimination.

# Results
## City-level Data

Table 4 gives the percentage of drivers who are black amongst those stopped in the light and in the darkness across all 7 cities. We see in Table 4 that, without controlling for any other variables, the proportion of minority drivers increases in the dark, suggesting the opposite of VOD discrimination.

```{r raw_prop_city}
data.cities %>% 
  group_by(is_dark) %>% 
  summarise(prop_minority = sum(is_minority_demographic) / n(), total = n()) %>% 
  mutate(
    is_dark = if_else(is_dark == TRUE, "Dark", "Light"),
    prop_minority = scales::percent(prop_minority)
  ) %>% 
  knitr::kable(caption = "Percentage of stopped drivers who are black across all citeis, by darkness", col.names = c("", "Percentage Minority Drivers", "Total # Stops"))
```

Figure 1 shows the percentage of stopped drivers who are black in the 90 minutes before and after sunset, controlling for clock time. Both clock time and time relative to sunset is binned by 15 minutes. The proportion is calculated across all cities and one proportion is calculated per combination of clock time and time relative to sunset. The restriction of stops to those between 6:30 PM and 7:30 PM ensures that there are stops in every city in each bin of time relative to sunset and each bin of clock time.

```{r full_city_plot}
plot_prop_minority_by_time(
  data.cities,
  title = "Figure 1",
  subtitle = "The percentage of stopped drivers who are black is lower in the dark,\nwhen controlling for clock time",
  min_clock_time = hms::hms(hours = 18, min = 30), 
  max_clock_time = hms::hms(hours = 19, min = 30),
  smooth_method = "lm"
)
```

Figure 1 suggests that, between 6:30 and 7:30 PM, the proportion of minority drivers in the population of stopped drivers is lower after dark than before dark. The percentage of drivers stopped between 90 and 85 minutes before sunset is between 0.5 and 3 percentage points higher than the percentage of minority drivers amongst those stopped between 90 and 105 minutes after sunset.[^4]

[^4]: It should be noted that if we calculate the proportion of minorities by city (or geographic subdivision) and plot the mean of these proportions for each clock time and time relative to sunset, we produce a plot with very slightly positively sloped lines. This is likely due to the increased influence of smaller cities, which, as we see in Figure 2, tend to show less evidence of discrimination. We felt that Figure 1 is more representative of the effect present in the regressions included in Table 5, as it effectively weights the cities by their size, however, this phenomenon is worth further investigation.

Table 5 shows the estimated coefficient on the darkness indicator variable in the three logistic regression models described above.

```{r full_city_reg}
read_rds(here::here("cache", "vod_full_results.Rds")) %>% 
  knitr::kable(
    caption = "Logistic Regression Estimates (All stops; City data)", 
    col.names = c("Controls", "Coefficient Estimate", "Standard Error")
  )
```

As we would expect from Figure 1, the coefficient on the darkness variable is negative in all three models. The additional controls for geography and month of the year negligibly impact the estimate. All three coefficients are statistically significant at the 0.05 level, suggesting there may be VOD discrimination present. As G & R emphasise, the test provides only a qualitative measure of discrimination.

It is highly possible that discriminatory behavior varies by city. To explore this, we additionally fit the model that controls for clock time and geography on each city individually.  Figure 2 displays the estimated coefficients on the darkness variable with 95% confidence intervals for each city. Nashville, Philadelphia, San Francisco and Madison have negative, statistically significant coefficients, indicating that there is evidence of discrimination at the 0.05 level.

```{r by_city_full}
by_city <- read_rds(here::here("cache", "vod_geo_adj_by_city.Rds"))

by_city %>%  
  mutate(
    low = estimate - 1.96*std.error, 
    high = estimate + 1.96*std.error
  ) %>% 
  ggplot(aes(reorder(city, estimate), estimate)) +
  geom_hline(yintercept = 0, color = "red") +
  geom_point() +
  geom_errorbar(aes(ymin = low, ymax = high)) +
  theme(
    axis.text.x = element_text(hjust = 1, angle = 30)
  ) +
  labs(
    x = NULL,
    y = "Coefficient on Darkness Variable with 95% C.I.",
    title = "Figure 2",
    subtitle = "Clock-time, month and subgeography adjusted models show statistically significant\nVOD discrimination in Madison, San Francisco, Philadelphia and Nashville"
  )
```

Notice that the cities with positive coefficients, Arlington and Columbus, are smaller than cities with negative coefficients like Nashville, Philadelphia and San Francisco.  The largest estimate for the coefficient on the darkness variable is in Madison, WI. Figure 3 replicates Figure 1 using only the data from Madison. The highly downward sloping lines support the evidence for VOD discrimination found in the logistic regression. 

```{r madison_plot}
plot_prop_minority_by_time(
    data.cities,
    city_only = "Madison",
    title = "Figure 3",
    subtitle = "The percentage of stopped drivers in Madison who are black is lower in the dark,\nwhen controlling for clock time",
    min_clock_time = hms::hms(hours = 18, min = 30), 
    max_clock_time = hms::hms(hours = 19, min = 30),
    smooth_method = "lm"
  )
```

### DST Shift
We now limit our analysis to the two weeks either side of the two DST shifts each year. Since the clock time changes by one hour at each shift, there is approximately one hour of the day during these four week periods where it is light in the first two weeks and dark in the second two weeks, or vice versa. To generate Figure 4, we limit the data to stops that occur in the 20 minute period in the middle of this one-hour discontinuity. We aggregate the data over cities and years by day relative to the DST shift. Figure 4 shows the percentage of drivers stopped in this period who are minority drivers.

```{r discontinuity_city}
date_lims <- 
  data.cities %>% 
  filter(is_dst_period == TRUE) %>% 
  summarise(min = min(date), max = max(date))

city_years <- 
  data.cities %>% 
  distinct(city, year = year(date))

city_centroids <-
  data.cities %>% 
  group_by(city) %>% 
  summarise(
    lat = median(lat),
    lng = median(lng),
    tz = min(tz)
  )

dst_sunsets <-
  tibble(
     date = seq(date_lims$min, date_lims$max, by = "days")
  ) %>% 
  mutate(
    is_dst = dst(as.POSIXct(date, tz = Sys.timezone())),
    is_dst_tomorrow = lead(is_dst),
    day_of_change = is_dst == FALSE & is_dst_tomorrow == TRUE | 
      is_dst == TRUE & is_dst_tomorrow == FALSE,
    day_before_change = lead(day_of_change) == TRUE
  ) %>% 
  filter(day_of_change == TRUE | day_before_change == TRUE) %>% 
  mutate(
    year = year(date),
    season = if_else(month(date) < 6, "fall", "spring")
  ) %>% 
  select(year, season, date, day_before_change, day_of_change) %>% 
  left_join(city_years, by = "year") %>% 
  left_join(city_centroids, by = "city") %>% 
  mutate(subgeography = city) %>% 
  left_join(infer_sunset_times(.), by = c("date", "tz", "subgeography"))

dst_times <-
  dst_sunsets %>% 
  mutate(
    day = if_else(day_before_change == TRUE, "day_before", "day_of"),
    sunset_minute = hour(hms(sunset)) * 60 + minute(hms(sunset))
  ) %>% 
  select(-day_before_change, -day_of_change, -date, -sunset, -dusk) %>% 
  spread(key = day, value = sunset_minute) %>% 
  mutate(
    earlier = if_else(day_before < day_of, day_before, day_of),
    later = if_else(day_before > day_of, day_before, day_of),
    min_minute = earlier + 20,
    max_minute = earlier + 40
  ) %>% 
  filter(max_minute < later) %>% 
  select(year, season, city, min_minute)

dst_dates <-
  dst_sunsets %>% 
  filter(day_of_change) %>% 
  distinct(year, season, date) %>% 
  select(year, season, dst_date = date)

full_data <-
  data.cities %>% 
  mutate(
    year = year(date),
    season = if_else(month < 6, "fall", "spring")
  ) %>% 
  left_join(dst_dates, by = c("year", "season")) %>% 
  left_join(dst_times, by = c("year", "season", "city")) %>% 
  mutate(days_since_dst = date - dst_date)

plot_data <-
  full_data %>% 
  filter(
    days_since_dst < days(15),
    days_since_dst > days(-15),
    minute >= min_minute,
    minute <= min_minute + 20
  ) %>% 
  mutate(
    is_dark = factor(is_dark, levels = c(TRUE, FALSE), labels = c("Dark", "Light")),
    season = str_to_title(season),
    before_dst = days_since_dst < 0
  ) %>% 
  group_by(days_since_dst, before_dst, season, is_dark, geo_control) %>% 
  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n()) %>%
  summarise(weighted_mean_prop = weighted.mean(prop_minority, n))
  
plot_data %>% 
  ggplot(aes(days_since_dst, weighted_mean_prop)) +
  geom_vline(xintercept = days(0), color = "grey60") +
  geom_point(aes(color = is_dark)) +
  geom_smooth(method = "lm", color = "grey40", se = FALSE) +
  scale_color_manual(values = c("#2b8cbe", "#feb24c")) +
  labs(
    title = "Figure 4",
    subtitle = "The trend in the proportion of minority drivers stopped in the 20 minute period\nat the center of the DST discontinuity suggests possible VOD discrimination",
    x = "Days since DST change",
    y = "Proportion minority drivers stopped",
    color = NULL
  ) +
  facet_grid(season ~ .)
```

The lines in Figure 4 are sloping slightly downwards in the direction of the days where it is dark at the time of the stops. This suggests there may be VOD discrimination apparent in the data.[^5]

[^5]: See 4. The same argument applies in this case.

From Figure 4, we expect to see negative coefficient estimates in the logistic regressions. Table 6 shows the regression estimates of the coefficient on the darkness variable with standard errors, for the simple model and the model that adjusts for geographic subdivision. The model that controls for month is not included, since the limitation to the DST period effectively controls for month.

```{r dst_reg_table_city, warning=FALSE}
dst_fall <- read_rds(here::here("cache", "vod_dst_fall.Rds"))
dst_spring <- read_rds(here::here("cache", "vod_dst_spring.Rds"))

tribble(
  ~season, ~adjustments, ~estimate,  ~se, 
  "Fall", "Clock time", coef(dst_fall$model_time_const)[2], coef(summary(dst_fall$model_time_const))[2, 2], 
  "", "Clock time and Geography", coef(dst_fall$model_geo_adjusted)[2], coef(summary(dst_fall$model_geo_adjusted))[2,2],
  "Spring", "Clock time", coef(dst_spring$model_time_const)[2], coef(summary(dst_spring$model_time_const))[2, 2],
  " ", "Clock time and Geography",  coef(dst_spring$model_geo_adjusted)[2], coef(summary(dst_spring$model_geo_adjusted))[2,2]
) %>% 
  column_to_rownames("season") %>% 
  knitr::kable(
    caption = "Logistic Regression Estimates (DST period; City data)",
    col.names = c("Controls", "Coefficient", "Standard Error"), 
    row.names = TRUE
  )
```

The coefficient estimates are all negative, however the estimates from the models that control for both clock time and geography are not statistically significant at the 0.05 level. 

Again, we may be interested in the city-by-city effects. Figure 5 shows the coefficient estimates for the geography adjusted model fit on the DST period data in each city and season with 95% confidence intervals. Though the trends are similar to those seen in the city-by-city analysis of the entire dataset, none of the coefficient estimates are statistically significant at the 0.05 level except for the Fall estimate for Philadelphia.

```{r by_city_dst}
by_city_dst <- 
  read_rds(here::here("cache", "vod_dst_by_city.Rds"))

by_city_dst %>%  
  mutate(
    low = estimate - 1.96*std.error, 
    high = estimate + 1.96*std.error,
    season = str_to_title(season)
  ) %>% 
  ggplot(aes(reorder(city, estimate), estimate)) +
  geom_hline(yintercept = 0, color = "red") +
  geom_point() +
  geom_errorbar(aes(ymin = low, ymax = high)) +
  theme(
    axis.text.x = element_text(hjust = 1, angle = 30)
  ) +
  labs(
    x = NULL,
    y = "Coefficient Estimate with 95% CI",
    title = "Figure 5",
    subtitle = "The clock-time and geography adjusted models fit on the data in the two weeks\neither side of the DST shift do not show statistically significant effects at the 0.05 level for any city."
  ) +
  facet_grid(season ~ .)
```

## Statewide Analysis

We now turn our attention to the statewide data. Table 7 displays the raw percentage of stopped drivers who are minority drivers during the light and during the darkness. The percentage differs by less than 0.01 percent, showing no indication of discrimination.

```{r raw_prop_state}
data.states %>% 
  group_by(is_dark) %>% 
  summarise(prop_minority = sum(is_minority_demographic) / n(), total = n()) %>% 
  mutate(
    is_dark = if_else(is_dark == TRUE, "Dark", "Light"),
    prop_minority = scales::percent(prop_minority)
  ) %>% 
  knitr::kable(
    caption = "Proportion of stopped drivers who are black across all states, by darkness", 
    col.names = c("", "Percentage Minority Drivers", "Total # Stops")
  )
```

Figure 6 replicates Figure 1 for the statewide data. We can see that the lines are negatively sloped. Between 6:30PM and 7:30PM, the percentage of drivers amongst those stopped 90-85 minutes before sunset is between 1.5 and 2.5 percentage points higher than the percentage of minority drivers amongst those stopped 90-105 minutes after sunset. From this figure we expect to see evidence of discrimination in the logistic regressions.

```{r statewide_plot}
plot_prop_minority_by_time(
  data.states,
  title = "Figure 6: Statewide Proportion of Minority Drivers By Time",
  subtitle = "The percentage of stopped drivers who are black is lower in the dark,\nwhen controlling for clock time",
  min_clock_time = hms::hms(hours = 18, min = 30), 
  max_clock_time = hms::hms(hours = 19, min = 30),
  smooth_method = "lm"
)
```

The results of the logistic regression models controlling for clock time only and for clock time and geography are shown in Table 8. The coefficient on the darkness variable is negative in both models, and larger in magnitude in the model that controls additionally for geography. Both coefficient estimates are statistically significant at the 0.05 level, suggesting there is VOD discrimination present.

```{r statewide_models}
statewide_models <- 
  list(
    time_only = read_rds(here::here("cache", "vod_statewide_time_const.rds")),
    time_geo = read_rds(here::here("cache", "vod_statewide_geo_adj.rds"))
  )

tribble(
  ~controls, ~estimate,                                    ~se,
  "Clock time", statewide_models$time_only$estimate[2], statewide_models$time_only$std.error[2],
  "Clock time and Geography", statewide_models$time_geo$estimate[2], statewide_models$time_geo$std.error[2]
) %>% 
  knitr::kable(
    caption = "Logistic Regression Estimates (All stops; Statewide data)", 
    col.names = c("Controls", "Coefficient Estimate", "Standard Error")
  )
```

#### TODO: State by state plot & maybe discontinuity

# Discussion
With the exception of the clock time and geography adjusted models fit on only the city data in the DST period, we see statistically significant evidence of VOD discrimination across the logistic regression models fit on both the city-level and the statewide data. When we consider discrimination at a city-level, we see statistically significant discrimination in Madison, San Francisco, Philadelphia and Nashville. At the state-level __TODO__.

The lack of statistical significance in the DST analysis when adjusting for geographic subdivision is likely indicative that the restriction of the data to only four weeks of the year significantly reduces the power of the test. The coefficient estimates in the models fit only on the DST shift data look qualitatively similar to those in the models fit on all of the data, and the model fit on all the data with a control for month showed no significant difference from those that do not control for month. This provides further reason to believe that the discrepancy in statistical significance is a symptom of the significant reduction in the size of the datasets. The quantity of data could be increased by including cities that do not have subdivision data available and running only the clock-time adjusted regression.[^6] It will also be interesting to see whether the DST analysis on the much larger statewide dataset has the same issues of power.

[^6]: See 1 for the list of these cities.

Despite the high statistical significance observed in many of the models, there are valid concerns with the premise of the veil of darkness test that should be considered. Firstly, the test makes the assumption that darkness is a good proxy for the visibility of a drivers. Further, it assumes that darkness is a binary variable, which implicitly assumes that the "extent" of darkness is the same across all geographic subdivisions at all times after dusk. This ignores factors such as street and ambient lighting which we could justifiably expect to vary significantly between urban, suburban and rural cities and counties.

It also assumes that the visibility of a driver's race is the factor on which discrimination is based. As G & R point out, officers may be using vehicle characteristics to infer a driver's race, a practice called "car profiling", rather than explicitly observing of the driver. In addition, it is feasible that discriminatory behavior varies significantly by officer, which is not accounted for. While our analysis controls for other behaviors that may impact the proportion of minority drivers stopped, like disporportionately patrolling in heavily minority neighborhoods or at times when more minority drivers are driving, through clock time and geographic controls, behaviors like "car profiling" or officer-level effects are not addressed.
