---
title: "A large-scale analysis of racial disparities in police stops across the
United States"
header-includes:
  - \usepackage{pdflscape}
  - \usepackage{graphicx}
---

```{r setup, echo = FALSE}
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

pfs <- read_rds(here::here("results", "prima_facie_stats.rds"))
vod <- read_rds(here::here("results", "veil_of_darkness.rds"))
disp <- read_rds(here::here("results", "disparity.rds"))
mj <- read_rds(here::here("results", "marijuana_legalization_analysis.rds"))
```

\begin{center}
\section{Basic Numbers}
\end{center}

We have collected and released a dataset of 
`r pfs$counts$collected$n_stops` traffic stops carried out by 
`r pfs$counts$collected$n_states` state patrol agencies and 
`r pfs$counts$collectec$n_cities` municipal police departments.

We analyzed a unique dataset detailing 
`r pfs$counts$analyzed$n_stops` traffic stops carried out by 
`r pfs$counts$analyzed$n_states` state patrol agencies and 
`r pfs$counts$analyzed$n_cities` municipal police departments.

\begin{center}
\section{Prima Facie Stats}
\end{center}

```{r stop_rates}
pfs$plots$stop
pfs$rates$stop %>% 
  group_by(subject_race) %>%
  summarize(average_annual_stop_rate = mean(annual_stop_rate))
  kable()
```

```{r search_rates}
pfs$plots$search
pfs$rates$search %>% 
  group_by(subject_race) %>%
  summarize(average_search_rate = mean(search_rate))
  kable()
```


\newpage
\begin{center}
\section{Veil of Darkness}
\end{center}

```{r veil_of_darkness-plot}
vod$full$plots$states$time_sliced$`Statewide, TX`
```

```{r veil_of_darkness-mod}
bind_rows(
  vod$dst$coefficients %>% 
    mutate(model = "DST"),  
  vod$full$coefficients %>% 
    mutate(model = "Full year")
) %>% 
  select(model, is_dark, std_error, everything()) %>% 
  kable()
```


\newpage

\begin{center}
\section{Search Disparity Analysis}
\end{center}

\begin{center}
\subsection{Outcome test results}
\end{center}

```{r disparity-city-hit_rates-table}
kable(disp$city$outcome$results$aggregate_hit_rates)
```
```{r disparity-state-hit_rates-table}
kable(disp$state$outcome$results$aggregate_hit_rates)
```
\newpage
```{r disparity-city-hit_rates-plot, fig.height = 3, fig.width = 7}
disp$city$outcome$plots$aggregate
```
```{r disparity-state-hit_rates-plot, fig.height = 3, fig.width = 7}
disp$state$outcome$plots$aggregate
```

\begin{center}
\subsection{Threshold test results}
\end{center}
```{r disparity-city-thresholds-table}
kable(disp$city$threshold$results$aggregate_thresholds)
```
```{r disparity-state-thresholds-table}
kable(disp$state$threshold$results$aggregate_thresholds)
```
\newpage
```{r disparity-city-thresholds-plot, fig.height = 3, fig.width = 7}
disp$city$threshold$plots$aggregate
```
```{r disparity-state-thresholds-plot, fig.height = 3, fig.width = 7}
disp$state$threshold$plots$aggregate
```

\newpage
\begin{center}
\section{Marijuana Legalization Analysis}
\end{center}
```{r mj_table}
kable(mj$tables$search_rate_difference_in_difference_coefficients)
```

\newpage
```{r mj_test_search_rates_plot, fig.height = 3, fig.width = 7}
mj$plots$test_search_rates +
  theme(
    legend.position = c(0.88, 0.80),
    axis.title.y = element_text(size = 12)
  )
```


```{r mj_test_misdemeanor_rates_plot, fig.height = 3, fig.width = 7}
mj$plots$test_misdemeanor_rates +
  theme(
    legend.position = c(0.88, 0.80),
    axis.title.y = element_text(size = 12)
  )
```


```{r mj_inferred_threshold_plot, fig.height = 3, fig.width = 7}
mj$plots$inferred_threshold_changes
```


```{r save_mj_control_search_rates}
# NOTE: have to save the file to the tmp directory because landscape doesn't
# work with r execution blocks
library(ggplot2)
ggsave(
  "/tmp/mj_control_search_rates.png",
  mj$plots$control_search_rates,
  height = 6,
  width = 11
)
```
\begin{landscape}
\includegraphics{/tmp/mj_control_search_rates.png}
\end{landscape}
